Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO campaign

### 2022-05-10 analysis/paper/paper_ana_20181122.tex
comentarios de Maru

  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble. We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND, and satellite radiances from different microwave and infrared sensors. We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation. We also run ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation. We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
  \end{abstract}
   \begin{keyword} Regional Data Assimilation, Surface Obsevations, Satellite Observations\end{keyword}
 \end{frontmatter}

@@

In that regard, several authors investigated the impact of assimilating surface weather data (e.g.~Wheatley and Stensrud (2010), Ha and Snyder (2014), Chang et al. (2017), Bae and Min (2022), Banos et al. (2021), Maejima et al. (2019), and Chen et al. (2016) using different assimilation methodologies. Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems. Sobash and Stensrud (2015) showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours). More recently, Gasperoni et al. (2018) evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis. They found a positive effect of these observations upon the initiation of deep moist convection. This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.

@@

We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis. with random perturbations to generate the initial ensemble perturbations. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5\(^{\circ}\) horizontal grid spacing with a smooth time evolution as in Necker et al. (2020) and Maldonado et al. (2021). In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. This method helps to prevent an underestimation of the ensemble spread (Ouaraini et al., 2015), and the random perturbations are also applied at the boundaries to maintain proper levels of ensemble spread within the domain.

@@

The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream. We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP). These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft. The orange triangles in Figure \ref{fig:dominio} indicate the location of the surface stations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes. Wind surface observations over oceans (ASCATW) come from scatterometers and are also included in the PREPBUFR files.

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations, are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2). Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. In particular, the multispectral sensors provided between 100 and 1000 observations for every scan every 12 hours, contributing 88 \% of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere.
The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa). The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. First, a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred to those over land or snow (Hu et al., 2018).

@@

In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves for the \(\beta_i\) that provides the best fit between the simulation and the observations. The coefficients were initialized at 18 UTC Nov 18, 2018 with the GFS system coefficients. The assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each time.

@@

where \(O\) and \(X\) stand for independent observations, and the simulations respectively and N is the sample size.

For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, Roberts (2008)) for different neighborhood length scales and thresholds. Following Maldonado et al. (2021) the definition of the FSS is slightly modified in that the probability of having rain exceeding a particular threshold and over a particular region is computed taking into account all the ensemble members :

\[\mathit{FSS} = 1-\frac{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i)^{2}+\frac{1}{N}\sum_{i=1}^{N} ({P_o}_i)^{2}} \]

where \(P_{xi}\) is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and \(P_{oi}\) is the proportion of grid points in each sampling area in which the observed accumulated precipitation is over the same threshold. Since the precipitation forecast of all the ensemble members is considered in the computation of \(P_{xi}\), the smoothing effect of the ensemble mean is avoided. The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.

@@

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the \(mean RCRV\) and the \(sd RCRV\) for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation. Figure 5 shows the \(sd RCRV\) for surface observations box-averaged to a 2.5° grid. The \(sd RCRV\) for wind observations (Figure \ref{fig:rcrv-sfc}a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error. For the temperature (Figure \ref{fig:rcrv-sfc}b), the results are similar except that for some areas in the west of the domain the \(sd RCRV\) can be as high as 4.5. In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.

@@

The analyzed distribution of temperature and moisture in the PBL (Figure \ref{fig:summary-fields}b) resembles the characteristics observed in the temperature profiles (Figure \ref{fig:TQ-diff}a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer that SATWND. On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations. Figure \ref{fig:summary-fields}c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 \(km\) wind shear. The values of MCAPE in CONV do not exceed 2000 \(JKg^{-1}\) while the rest of the experiments show maximum MCAPE over 4000 \(JKg^{-1}\). MCAPE in the RAD experiment is lower compared to AWS or SATWND. This is consistent with less humidity in the PBL with respect to these experiments but may be partially compensated by a slightly warmer PBL in the RAD experiment. The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 \(ms^{-1}\) at the southern tip of the region with positive MCAPE values. Moreover, in this same region, these experiments show larger MCAPE values than CONV. Note that wind shear over 15 \(ms^{-1}\) is associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowski and Richardson, 2010).

@@

First, we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation. Figure \ref{fig:pp-hov}a shows the mean hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) (Clark, 2017) for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments. The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \ref{fig:pp-hov}a). In all the experiments, the accumulated precipitation in the short-range forecasts is underestimated. This is particularly evident in CONV (Figure \ref{fig:pp-hov}b), where the convection initiation is delayed and occurs further north with respect to the observed initiation. AWS, SATWND, and RAD better capture the timing and location of convective initiation (Figures \ref{fig:pp-hov}c-e). AWS and RAD show a more fragmented distribution compared with SATWND, possibly due to the development of less organized convection during Nov 22. After 18 UTC Nov 22, RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of enhanced development of the convection.

@@

To quantify the spatial match between the observed precipitation and the short-range forecast precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}). All experiments show similar values of FSS during the initialization of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period. This indicates that radiance observations have a positive impact on the analysis. The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS. AWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for this case study. The assimilation of radiances led to an overall improvement of the 1-hour forecast precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov 22 (Figure \ref{fig:fss}b,d). The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and also for spatial scales above 500 km, not shown).

@@

\begin{figure*}[h]
\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecast ensemble mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018. Black circles in first row show the observation range of each radar.}\label{fig:dbz-mean}
\end{figure*}

For IOP 8 (Figures \ref{fig:soundings}e-h), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulation. This area was also behind the cold front and affected by low-level cold advection. The assimilation of AWS, SATWND, and RAD reduces the cold bias and RMSE for temperature between 5 and 12 km and the RMSE in the PBL compared with CONV (Figure \ref{fig:soundings}e). The reduction of bias and RMSE is also important for dew point temperature (Figure \ref{fig:soundings}f) with SATWND showing the bigger impact followed by AWS and RAD. The zonal wind is overestimated in the analyses and only RAD shows an improvement with respect to CONV in the upper troposphere (Figure \ref{fig:soundings}g). At low levels the meridional wind (Figure \ref{fig:soundings}g) presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AWS, SATWND, and RAD. In fact, low level biases in these experiments are higher than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of AWS).

@@

In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment. We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the 1-h forecast to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}). CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations. AWS, SATWND, and RAD show improvements in the FSS values, particularly for the higher threshold (Figure \ref{fig:fssfcst}b, d). Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD tham the forecast initialized at 00 UTC, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.

@@

To summarize, in this study we conclude that the assimilation of surface observations with higher spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation. Moreover, the ensemble forecasts initialized from the analysis show promising results to predict the precipitation of severe events. In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like GPS radio occultation data and radiances from geostationary orbiting satellites.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

>>>>

  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble. We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors. We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation. We also ran ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation. We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
  \end{abstract}
   \begin{keyword} Regional Data Assimilation, Surface Obsevations, Satellite Observations\end{keyword}
 \end{frontmatter}

@@

In that regard, several authors investigated the impact of assimilating surface weather data (e.g.~Wheatley and Stensrud (2010), Ha and Snyder (2014), Chang et al. (2017), Bae and Min (2022), Banos et al. (2021), Maejima et al. (2019), and Chen et al. (2016)) using different assimilation methodologies. Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems. Sobash and Stensrud (2015) showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours). More recently, Gasperoni et al. (2018) evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis. They found a positive effect of these observations upon the initiation of deep moist convection. This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.

@@

We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis with random perturbations to generate the initial ensemble perturbations. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5\(^{\circ}\) horizontal grid spacing with a smooth time evolution as in Necker et al. (2020) and Maldonado et al. (2021). In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. This method helps to prevent an underestimation of the ensemble spread (Ouaraini et al., 2015), and the random perturbations are also applied at the boundaries to maintain proper levels of ensemble spread within the domain.

@@

The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream. We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP). These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft. The orange triangles in Figure \ref{fig:dominio}a indicate the location of the surface stations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes. Wind surface observations over oceans (ASCATW) come from scatterometers and are also included in the PREPBUFR files.

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations, are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \ref{tab:table-rad}). Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. In particular, the multispectral sensors provided between 100 and 1000 observations for every scan every 12 hours, contributing 88 \% of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere.
The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa). The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. First, a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred to those over land or snow (Hu et al., 2018).

@@

In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves the \(\beta_i\) that provides the best fit between the simulation and the observations. The coefficients were initialized at 18 UTC Nov 18, 2018 with the GFS system coefficients. The assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each time.

@@

where \(O\) and \(X\) stand for independent observations and the simulations respectively, and N is the sample size.

For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute the Fractions Skill Score (FSS, Roberts (2008)) for different neighborhood length scales and thresholds:

\[\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} \]

Where \(P_{oi}\) is the fraction of grid points in the \(i-th\) sampling area in which the observed accumulated precipitation is greater than a specified threshold. Following Maldonado et al. (2021), \(P_{xi}\) is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the \(i-th\) sampling area.
The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.

@@

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the \(mean RCRV\) and the \(sd RCRV\) for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation. Figure \ref{fig:rcrv-sfc} shows the \(sd RCRV\) for surface observations box-averaged to a 2.5° grid. The \(sd RCRV\) for wind observations (Figure \ref{fig:rcrv-sfc}a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error. For the temperature (Figure \ref{fig:rcrv-sfc}b), the results are similar except that for some areas in the west of the domain the \(sd RCRV\) can be as high as 4.5. In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.

@@

The analyzed distribution of temperature and moisture in the PBL (Figure \ref{fig:summary-fields}b) resembles the characteristics observed in the temperature profiles (Figure \ref{fig:TQ-diff}a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer than in SATWND. On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations. Figure \ref{fig:summary-fields}c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 \(km\) wind shear. The values of MCAPE in CONV do not exceed 2000 \(JKg^{-1}\) while the rest of the experiments show maximum MCAPE over 4000 \(JKg^{-1}\). MCAPE in the RAD experiment is lower compared to AWS or SATWND. This is consistent with less humidity in the PBL with respect to these experiments but may be partially compensated by a slightly warmer PBL in the RAD experiment. The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 \(ms^{-1}\) at the southern tip of the region with positive MCAPE values. Moreover, in this same region, these experiments show larger MCAPE values than CONV. Note that wind shear over 15 \(ms^{-1}\) is associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowski and Richardson, 2010).

@@

First, we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation. Figure \ref{fig:pp-hov}a shows the hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) (Clark, 2017) for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments. The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \ref{fig:pp-hov}a). In all the experiments, the accumulated precipitation in the short-range forecasts is underestimated. This is particularly evident in CONV (Figure \ref{fig:pp-hov}b), where the convection initiation is delayed and occurs further north with respect to the observed initiation. AWS, SATWND, and RAD better capture the timing and location of convective initiation (Figures \ref{fig:pp-hov}c-e). AWS and RAD show a more fragmented distribution compared with SATWND, possibly due to the development of less organized convection during Nov 22. After 18 UTC Nov 22, RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of enhanced development of the convection.

@@

To quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}). All experiments show similar values of FSS during the initiation of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period. This indicates that radiance observations have a positive impact on the analysis. The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS. AWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for this case study. The assimilation of radiances led to an overall improvement of the 1-hour forecast precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov 22 (Figure \ref{fig:fss}b,d). The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and also for spatial scales above 500 km, not shown).

@@

\begin{figure*}[h]
\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecast probability matched mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018. Black circles in first row show the observation range of each radar.}\label{fig:dbz-mean}
\end{figure*}

For IOP 8 (Figures \ref{fig:soundings}e-h), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulation. This area was also behind the cold front and affected by low-level cold advection. The assimilation of AWS, SATWND, and RAD reduces the cold bias and RMSE for temperature between 5 and 12 km and the RMSE in the PBL compared with CONV (Figure \ref{fig:soundings}e). The reduction of bias and RMSE is also important for dew point temperature (Figure \ref{fig:soundings}f) with SATWND showing the biggest impact followed by AWS and RAD. The zonal wind is overestimated in the analyses and only RAD shows an improvement with respect to CONV in the upper troposphere (Figure \ref{fig:soundings}g). At low levels the meridional wind (Figure \ref{fig:soundings}g) presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AWS, SATWND, and RAD. In fact, low level biases in these experiments are higher than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of AWS).

@@

In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23. We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}). CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations. AWS, SATWND, and RAD show improvements in the FSS values, particularly for the higher threshold (Figure \ref{fig:fssfcst}b, d). Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD than the forecast initialized at 00 UTC, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.

@@

To summarize, in this study we conclude that the assimilation of surface observations with high spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation. Moreover, the ensemble forecasts initialized from the analysis show promising results to predict the precipitation of severe events. In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like GPS radio occultation data and radiances from geostationary orbiting satellites.

\hypertarget{code-and-data-availability}{%
\section{Code and data availability}\label{code-and-data-availability}}

A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at \url{https://github.com/paocorrales/mesoda}. The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/{[}doi{]}, reference number {[}reference number{]}.

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}
